---
output:
  html_document: default
  pdf_document: default
params:
  pwd:
    label: "Please enter password to build connect to Postgres"
    value: ""
    input: password
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## EDA of NFL Data Bowl

Summary of Kaggle competition dataset for NFL Data Bowl. Further information is available at 
https://www.kaggle.com/c/nfl-big-data-bowl-2020 

```{r set_report, echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(RPostgreSQL)
library(ggthemes)
library(ggimage)

# Set up database connection
drv <- dbDriver("PostgreSQL")
conn <- dbConnect(drv, user ='postgres', host='localhost', dbname ='nfl_data_bowl', 
  password=params$pwd)

# Load data into workspace
nfl_data_raw <- dbGetQuery(conn, "select * from nfl_data")
```

```{r data_cleaning, echo=FALSE}
# make sure that appropriate transformations have taken place. 
nfl_data <- nfl_data_raw %>% 
  mutate(yards = as.numeric(yards),
         x = as.numeric(x),
         y = as.numeric(y),
         a = as.numeric(a),
         dis = as.numeric(dis),
         s = as.numeric(s))
```

The dataset contain data from `r n_distinct(nfl_data$gameid)` NFL games which took place between the `r min(nfl_data$season)` and `r max(nfl_data$season)` seasons. Across the entire dataset, the average yardage gained is `r round(mean(nfl_data$yards),1)` per play. The plot below breaks this down between teams. 

```{r avg_rushing_yards, echo=FALSE}

# Get NFL team logos
file <- list.files("/Users/samtaylor/Documents/Git/nfl_data_bowl/documentation/nfl_logos")

logos <-  list()
for (logo in 1:length(file)) {
  logos[logo] <- paste0("/Users/samtaylor/Documents/Git/nfl_data_bowl/documentation/nfl_logos/", file[logo])
}

# Get avg rushing yards and combine with logos
team_yards <- nfl_data %>% 
  group_by(possessionteam) %>% 
  summarise(avg_ypc = mean(yards)) %>% 
  mutate(team = fct_reorder(possessionteam, avg_ypc)) %>% 
  cbind(data.frame(logo = unlist(logos)))

ggplot(team_yards, aes(team, avg_ypc)) +  geom_col(alpha=3/4, color='black') +
  geom_image(aes(image=logo)) + theme_fivethirtyeight() +
  theme(axis.title = element_text()) + ylab('Average Yards Per Carry') +
  scale_x_discrete() + coord_flip() + geom_hline(yintercept = mean(nfl_data$yards), color='red', lty=2) + 
  xlab('') + ggtitle("Which team has the most yards/play since 2017?")

```

```{r lm_model}
play_data <- nfl_data %>% 
  group_by(playid, team) %>% 
  summarise(mean_speed = mean(s, na.rm = T),
            mean_acc = mean(a, na.rm = T),
            mean_dis = mean(dis, na.rm = T),
            yards = min(yards)) %>% 
  gather(key, value, -playid, -yards, -team) %>% 
  unite("key", team, key, sep='_') %>% 
  spread(key, value) %>% 
  ungroup()

lm(yards~., data=sample_play_data)
```


```{r}
ind <- sample(play_data, size =0.5)

```

```{r}
nfl_data %>% 
  ggplot(aes(abs(yards))) + geom_histogram(binwidth = 1)

marginal_probs = nfl_data %>% 
  mutate(yards = round(yards,0)) %>% 
  count(yards) %>% 
  mutate(total=sum(n),
         prob = n/total)

data <- data.frame(yards = c(-99:99)) %>% 
  left_join(marginal_probs) %>% 
  complete(prob, fill = list(prob=0)) %>% 
  arrange(yards) %>% 
  select(prob, yards)
  
samp <- sample(data$yards, 1000000, prob=data$prob, replace=T) 


   data.frame(samp) %>% 
     ggplot(aes(x=samp)) + geom_histogram(binwidth = 1)


```

```{r}
nfl_data %>% 
  filter(playid =='20170907000118') %>% 
  ggplot(aes(x,y, color=position)) + geom_point() + xlim(0,120) + ylim(0,53) +
  geom_vline(xintercept=74)  
```


